A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-12 docs/icccn2009/paper.tex
ICCCN 2009 Paper update, sections Message, Matcher, Broker, YSI sonde
- Section Message has updates regarding the types of 
  messages, and others;
  
- Adding images of XML schema to clean up a bit the
  presentation of the paper. This is just an experimental
  version, and can be replaced back; YSI XML Schema
  code was replaced by the XML Schema type image;
  
- Creating the match_config.xsd, the XML Schema that
  validates the Matcher rules. Instead of using code,
  adding the XML Schema type image for Matching Rule.
 
- Adding a DSP Broker section, since it's one critical
  component for

\begin{figure}
\centering
\epsfig{file=dsp, width=7cm}
\caption{\label{FIG_DSP} Data Sensor Platform (DSP).}
\end{figure}

@@

Message passing is the main paradigm for exchanging information within
a DSP. A data producer sends a message with data consumers accept
messages. We make use of XML technologies to define the format of a
message. Specifically, the common header of all messages are defined
via the following XML Schema:

\begin{code}
<xs:schema ...>
  <xs:complexType name="DSPMessage">
    <xs:element name="Header" type="Header"/>
  </xs:complexType>

  <xs:complexType name="Header">
    <xs:sequence>
      <xs:attribute name="messageID"/>
      <xs:attribute name="originatingDP"/>
      <!-- ... -->
    </xs:sequence>
  </xs:complexType>
</xs:schema>
\end{code}

The above XML Schema DSPMessage serves as a base schema for all
messages and itself defines information typically contained in the
header of a protocol data unit. Amongst others, the header contains a
unique message ID (line 8) and the originating DP (line 9). Using the
XML Schema, we use JAXB to generate a type safe API for handling
instances of DSPMessage. Different messages with different payloads
can be defined by creating schemas that derive from DSPMessage. While
inside the DSP all messages are of type DSPMessage, specific DP and DC
can distinguish between different message types based on their runtime
type information.

@@

can only route messages within one DSP. No notion of remote DSP
components.

@@

Our implementation of the DSP is based on the Open Source
implementation of OSGi called Knopflerfish \cite{knopflerfish01}. Since Sun
Microsystems does not support its JDK for the Gumstix, we use JamVM
\cite{jamvm01} as the Java virtual machine implementation and GNU Classpath
\cite{classpath01} for the runtime libraries. We have implemented various DSP
Components that implement the functionality needed for NetBEAMS. On
the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
case the aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
component that uses the RS232 standard to read the physical measurements from
the device over a serial port. This component converts the physical measurements 
into a DSP-specific message. The following XML Schema defines the YSI Sonde
specific message that captures all measurement points mentioned in
Table \ref{TAB_SONDE_MEASUREMENTS}:

\begin{code}
<xs:schema ...>
<xs:import schemaLocation="dsp-message.xsd"/>

<xs:element name="SondeDataMessage">
  <xs:complexType>
    <xs:complexContent>
      <xs:extension base="DSPMessage">
        <xs:sequence maxOccurs="unbounded"
                     minOccurs="1">
          <xs:element name="sondeData"
                      type="sondeDataType"/>
        </xs:sequence>
      </xs:extension>
    </xs:complexContent>
  </xs:complexType>
</xs:element>

<xs:complexType name="sondeDataType">
  <xs:sequence>
    <xs:element name="Date" type="xs:date"/>
    <xs:element name="Time" type="xs:time"/>
    <xs:element name="Temp" type="xs:float"/>
    <xs:element name="SpCond" type="xs:float"/>
    <!-- ... -->
  </xs:sequence>
</xs:complexType>
</xs:schema>
\end{code}

The schema is derived from DSPMessage. As a payload it adds an
unbounded sequence of sondeDataType. Each instance of this type
captures one timestamped measure point.

The YSI Sonde component can be
configured via special DSP messages that define properties such as the
sampling frequency. These property messages originate from the web
frontend that we will explain in a subsequent section.

>>>>

\begin{figure}
\centering
\epsfig{file=dsp-platform, width=7cm}
\caption{\label{FIG_DSP} Data Sensor Platform (DSP).}
\end{figure}

@@

Message passing is the main paradigm for exchanging information within
a DSP. After producing data, DP component wraps that data in the 
body of a message, so that DC components can receive the message and consume 
the data on the message's body. When it comes to passing messages to a remote
DSP instane, messages are serialized in XML \cite{xml2000} and automatically 
deserialized for DC located in the remote DSP. In order to serialize and
deserialize DSP messages, the DSP infrastructure uses the XML Schema 
standard \cite{xml-schema2004} to generate and validate the DSP message. Figure
\ref{FIG_DSP_MESSAGE} defines the structure of a general DSP Message.

\begin{figure}[!htb]
 \centering
 \epsfig{file=dsp-message, width=9cm, height=4cm}
 \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP Message.}
\end{figure}

The above XML Schema is defined for an ``Abstract Message'' type, whose
structure serves as a base schema for all different types of messages. Any
message will contain two basic attributes: ``messageID'' and ``ContentType''. 
The former uniquely identifies a message and the latter the data in the body of
the message. Complementary identification information such as the message 
producer and consumer are located in the header of the message, which are
defined by the schema type ``ComponenIdentifier''. A ``ComponentIdentifier''
carries out information about the component name and its physical host address.
Finally, the body of the message may contain any data type defined by the component 
developer, and it can be interpreted as the DSP payload.

Although XML schema is the de-facto standard to bind and validate xml data to
its type, the architectural infrastructure of DSP uses an XML Data Bindining
Technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}. It is used to generate
a type safe API for handling in-memory object graphs from the XML instances of 
DSP Messages and the content of its data payload. Moreover, taking advantage 
of the data type polimorphirsm of XML schema data types - concept borrowed from 
Object-Oriented Design \cite{Rumbaugh1991} - we defined different types of 
messages in order to allow specific DP and DC interpret the payload of any 
given message solely based on their runtime type, as follows:

\begin{itemize}
  \item Action Message;
  \item Event Message;
  \item Create Message;
  \item Insert Message;
  \item Update Message;
  \item Delete Message;
  \item Query Message;
  \item Measure Message.
\end{itemize}

Not only the message structure is defined by XML schemas, but also the body
content is and may be of any user-defined type. Therefore, the data
defined to be sent from any given DP to any DC can be also in XML format or
in-memory. An example illustrating a DSP message is given in the YSI section.

\subsection{Broker}

As soon as a DP produces data and wraps it up in a DSP Message, the DP only
interact with the DSP platform via the DSP Broker by making the call
``Broker.deliver(DSP Message)''. In this way, DSP broker is responsible for
delivering the produced message to DSP consumers DCs, as well as verifying
whether the delivery of the message is going to be done locally or remotely.

\begin{itemize}
 \item Local delivery of messages: it uses local object reference to deliver the
DSP message;
 \item Remote delivery of messages: it uses any DSP component that is
specialized in remote delivery defined as ``Gateway DSP Component''.
\end{itemize}

The selection of the DCs for a given message to where the Broker will deliver
messages or use a gateway component is determined by an object called Matcher.

@@

The Matcher is an object responsible to find the possible set of DCs for a given
DSP message, and its ``deterministic intelligence'' relies on matching rules
defined by the DSP administrator. In this way, a set of matching rules is define
in XML and stored at the DSP run-time directory, where each matching rule is
uniquely identified by a rule ID and determines the following (see
figure \ref{FIG_MATCHER_RULE}):

\begin{figure}[!htb]
 \epsfig{file=dsp-match-rule, width=9cm, height=4cm}
 \caption{\label{FIG_MATCHER_RULE} Definition of Matcher Rule.}
\end{figure}

\begin{itemize}
 \item Match Criteria: given the DSP message delivered to the Broker, which
component(s) produced that message and in which node? It defines the name of
the component type of the producer and its physical location (IP address);
 \item Match Target: given the DSP message, which component(s) are going
to consume the message, and in which node? It defines the name of the component
type of the consumer, and optionally defines a gateway component.
\end{itemize}

\subsection{Matcher replies to Broker}

When analyzing the matching rules, the Matcher only selects the rules that
satisfy the Matching Criteria and forwards them back to the Broker. Once the
broker has the matching rules, it determines whether the delivery is done
directly to a component defined on the Match Target, or if the message is
delivered to the optional gateway. We have defined special gateway DSP
components that is used to transport the DSP Messages, serialized in
XML, to any remote DSP instance.

@@

Our implementation of the DSP is based on the Open Source
implementation of OSGi called Knopflerfish \cite{knopflerfish01}. Since Sun
Microsystems does not support its JDK for the Gumstix, we use JamVM
\cite{jamvm01} as the Java virtual machine implementation and GNU Classpath
\cite{classpath01} for the runtime libraries. We have implemented various DSP
Components that implement the functionality needed for NetBEAMS. On
the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
case the aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
component that uses the RS232 standard to read the physical measurements from
the device over a serial port. This component converts the physical measurements 
into a DSP-specific message. The following XML Schema on figure \ref{FIG_DSP_YSI_DATA} 
defines the YSI Sonde's specific message format which includes each measurement data point, described in
Table \ref{TAB_SONDE_MEASUREMENTS}, and its type:

\begin{figure}[!htb]
\centering
\epsfig{file=dsp-message-body-ysi, width=8cm, height=7cm}
\caption{\label{FIG_DSP_YSI_DATA} Structure of the YSI Component Data.}
\end{figure}

The schema describes two Java classes, sondeDataType and SondeDataMessage, 
which are used to package the Sonde's incoming data into a DSP message for delivery 
to a remote consumer.  The SondeDataMessage class extends the AbstractMessageContent 
class by adding a new type to the message payload, SondeDataType.  Each time the Sonde
transmits its payload of measurement data during a sampling interval, the data is converted 
into a SondeDataType which is then encapsulated within SondeDataMessage and then sent 
to a remote consumer.

The YSI Sonde component can also be configured via special DSP messages, to direct the 
Sonde to change its sampling frequency.  These property messages originate from the web 
frontend that we will explained in a subsequent section.
